\LARGE \bf Autonomous Object Manipulation using a Soft Planar Grasping Manipulator

### 2015-10-16 abstract.tex
review comments by David implemented

This paper presents the development of an autonomous motion planning algorithm for a soft planar grasping manipulator capable of grasp-and-place operations by encapsulation
\dr{under uncertainty in the position and shape of the object.}
The gripping end of the soft manipulator is fabricated without weakening seams using lost-wax casting instead of the commonly used multi-layer lamination process.
The soft manipulation system can grasp randomly positioned objects within its reachable envelope and move them to a desired location without human intervention.
The autonomous planning system leverages the compliance and continuum bending of the soft grasping manipulator to achieve repeatable grasps in the presence of uncertainty.
A suite of experiments is presented that demonstrates the system's capabilities.

>>>>

This paper presents the development of an autonomous motion planning algorithm for a soft planar grasping manipulator capable of grasp-and-place operations by encapsulation
\dr{with uncertainty in the position and shape of the object.}
The end effector of the soft manipulator is fabricated in one piece without weakening seams using lost-wax casting instead of the commonly used multi-layer lamination process.
The soft manipulation system can grasp randomly positioned objects within its reachable envelope and move them to a desired location without human intervention.
The autonomous planning system leverages the compliance and continuum bending of the soft grasping manipulator to achieve repeatable grasps in the presence of uncertainty.
A suite of experiments is presented that demonstrates the system's capabilities.

### 2015-10-16 introduction.tex
review comments by David implemented

%High Level Vision
Soft robots exhibit continuum body motion, large scale deformation, high compliance and adjustable impedance  compared to traditional rigid-bodied robots with high impedance \cite{trivedi2008soft}. 
% whereas humans and animals have extremities with variable impedance.
%Soft robots won't necessarily replace existing rigid robot technologies, but complement it, because they can enable static and dynamic manipulations systems in nature are capable of.
%A further motivating factor for this work is that soft robots are inherently safer for interaction with humans and the environment due to their structural compliance and they can be lighter, one of the reasons why we research on that path.
%Fluidic robots have uses in scenarios where a classical robot would cause sparks and are dangerous. Why do we need accuracy if we have compliance. A 3DOF gripper is inherently not soft. A rigid gripper can not manipulate delicate objects without a lot of sensing.

@@

Our goal is to develop a soft planar fluid-powered gripper and a motion planning algorithm that leverages a soft morphology to robustly grasp, drag and place objects \dr{of unknown geometry}. 
%By utilizing softness in the design of our robot, we gain the ability to adapt to environmental variation \cite{mcmahan2006field}, harmlessly collide with the environment \cite{marchese2014whole}, and build resilience to unexpected external loads \cite{tolley2014resilient}; 
\rkk{In this paper we describe a planar soft robot manipulator we developed toward this goal.
We focus on the design, fabrication, control, and planning aspects of this soft robot.}

@@

%Our Approach
%To address these challenges, 
%We developed planning and control algorithms for a soft robotic arm with gripper.
The fluid-powered gripper at the end of the arm can grasp an object through an open-loop controlled bending motion, even if the gripper is positioned relatively inaccurately in relation to the object to be grasped.
%The gripper can conform due to his compliance to variations in object geometry, allowing it to successfully grasp.
The design of the gripper itself is inspired by the work of Polygerinos et al.~\cite{polygerinos2013towards}
The design is advantageous for grasping, because it exhibits high curvature, minimal radial expansion, and remains compliant during actuation. 
We can repeatably fabricate the gripper using a lost-wax casting process instead of the commonly used soft lithography technique.%, a multi-layer lamination process.
Without weakening seams, the gripper is not prone to de-lamination under high deformations. 
By abandoning the need for a lamination process, arbitrary shaped internal channels can be achieved.

@@

Several manufacturing processes for soft biomimetic robots were reviewed by Cho et al. \cite{cho2009review}. 
The vast majority of soft elastomer robots rely on the processes of soft lithography \cite{xia1998soft} and/or shape deposition manufacturing \cite{cham2002fast}. 
Especially noteworthy is the use of wax for fabricating jammable skin chambers, which stiffens by vacuuming\cite{steltz2009soft}. 
The gripper presented in this work also uses a lost-wax molding technique, but with a different type of actuation in mind: the obtained cavity structures are inflated to cause the gripper to bend.

%\rkk{Volder et al. \cite{volder2010pneumatic} provides a detailed review of both pneumatic and hydraulic microactuators.} 
There are several hardware examples for soft grippers described in recent literature, we will mainly focus on fluidic-based systems.
Deimel and Brock \cite{deimel2013compliant} developed a pneumatically actuated three-fingered hand made of reinforced silicone that is mounted to a hard robot and capable of grasping.
More recently, they have developed an anthropomorphic soft pneumatic hand capable of dexterous grasps, which is not mounted to a robot, but instead just held by a human \cite{deimel2014novel}. 
Ilievski et al. \cite{ilievski2011soft} create through a soft-lithography fabrication a pneumatic starfish-like gripper composed of an array of silicone chambers and a PDMS membrane. 
The gripper is hanging on a string and grasps objects like an egg or a mouse in an open-loop controlled manner. 
Stokes et al. \cite{Stokes2014hybrid} use a soft elastomer quadrupedal robot attached to a wheeled robot to grip and retrieve objects. 
A puncture resistant soft pneumatic gripper is developed by Shepherd et al. in \cite{shepherd2013soft}. 
An alternative to positive pressure actuated soft grippers is a robotic gripper that makes use of granular material jamming developed by Brown et al. and detailed in \cite{brown2010universal}.
\rkk{Ikuta and Suzuki \cite{ikuta2002safety} showed a multiple-segment micro-hydraulic actuator for entering blood vessels.}
The soft octopus-inspired arms developed in \cite{calisti2010study} and \cite{calisti2011octopus} are not fluidic powered, but instead use cables to pull rigid fixtures embedded within an elastomer body. 
The arms were capable of grasping objects like pens or screws.
A soft robotic tentacle developed in \cite{martinez2013robotic} was able to hold a flower and a horseshoe-shaped object.
The closest related soft pneumatic actuator design to our current work is the fast Pneu-net designs by Mosadegh et al. \cite{mosadegh2014pneumatic} and by Polygerinos et al. \cite{polygerinos2013towards}.
These finger-like actuators deform with minimal volume change and can bend to high curvatures.
None of the above described grippers were controlled autonomously to perform their tasks and accordingly no statement on repeatability of the autonomous execution was given.

%talk about control
\rkk{Simulation results of an online motion planner for a planar continuum manipulators were presented by Xiao and Vatcha\cite{xiao2010realtime}.}
\rkk{This work was extended by Li and Xiao \cite{li2015general} to present a more general formulation to constrained, continuum manipulation.}
%More generally, the literature has several examples of soft pneumatic elastomer robots. 
%Many of these robots use open-loop controllers. 
%There are soft fluid-powered rolling robots \cite{correll2010soft, onal2011soft, marchese2011soft}. 
%Also, there are soft-bodied robotic fish powered by pneumatics \cite{marchese2014autonomous} and hydraulics \cite{katzschmann2014hydraulic}. 
%There are soft legged pneumatic walkers \cite{shepherd2011multigait, tolley2014resilient}, explosive jumpers \cite{shepherd2013using}, and 
Marchese et al. \cite{marchese2014design}  \cite{marchese2014whole} demonstrated closed-loop position control of multi-segment soft planar fluidic elastomer manipulators in free and confined space. 
Positioning is achieved using forward and inverse kinematic models and real-time camera measurements. 
%The manipulator curvatures are controlled in real-time using a curvature-volume cascaded control law.
The previous version \rkk{was only suitable for inspection tasks by moving the manipulator through a constrained environment without object interaction or manipulation.}
%Previously, there were only single markers along the arm, which worked for a limited motion range, but was prone to fail if the markers go out of sight or another object is introduced to the environment. 
%Also, previously many trials were needed to successfully arrive at a final position.

\rkk{The work presented here builds on the previous system. We now combine two soft actuator types to leverage their strengths in a manipulation task. A new control method demonstrates the capabilities of this new soft manipulator to actually do autonomous manipulations under uncertainties.}
%The control of the arm is improved from the previous version by using a more robust vision tracking approach.

>>>>

%High Level Vision
Soft robots exhibit continuum body motion, large scale deformation, high compliance, and adjustable impedance  compared to traditional rigid-bodied robots with high impedance \cite{trivedi2008soft}. 
% whereas humans and animals have extremities with variable impedance.
%Soft robots won't necessarily replace existing rigid robot technologies, but complement it, because they can enable static and dynamic manipulations systems in nature are capable of.
%A further motivating factor for this work is that soft robots are inherently safer for interaction with humans and the environment due to their structural compliance and they can be lighter, one of the reasons why we research on that path.
%Fluidic robots have uses in scenarios where a classical robot would cause sparks and are dangerous. Why do we need accuracy if we have compliance. A 3DOF gripper is inherently not soft. A rigid gripper can not manipulate delicate objects without a lot of sensing.

@@

Our goal is to develop a soft planar fluid-powered gripper and a motion planning algorithm that leverages a soft morphology to robustly grasp, drag, and place objects \dr{of unknown geometry}. 
%By utilizing softness in the design of our robot, we gain the ability to adapt to environmental variation \cite{mcmahan2006field}, harmlessly collide with the environment \cite{marchese2014whole}, and build resilience to unexpected external loads \cite{tolley2014resilient}; 
\rkk{In this paper we describe a planar soft robot manipulator we developed toward this goal.
We focus on the design, fabrication, control, and planning aspects of this soft robot.}

@@

%Our Approach
%To address these challenges, 
%We developed planning and control algorithms for a soft robotic arm with gripper.
The fluid-powered gripper at the end of the arm can grasp an object through an open-loop controlled bending motion, even if the gripper is positioned relatively inaccurately in relation to the object to be grasped.
%The gripper can conform due to his compliance to variations in object geometry, allowing it to successfully grasp.
The design of the gripper itself is inspired by the work of Polygerinos et al.~\cite{polygerinos2013towards}
The design is advantageous for grasping, because it exhibits high curvature, minimal radial expansion, and remains compliant during actuation. 
We can repeatably fabricate the gripper using a lost-wax casting process instead of the commonly used soft lithography technique.%, a multi-layer lamination process.
As a homogeneous piece without weakening seams, the gripper is not prone to de-lamination under high deformations. 
By abandoning the need for a lamination process, arbitrary shaped internal channels can be achieved.

@@

Several manufacturing processes for soft biomimetic robots were reviewed by Cho et al. \cite{cho2009review}. 
The vast majority of soft elastomer robots rely on the processes of soft lithography \cite{xia1998soft} and/or shape deposition manufacturing \cite{cham2002fast}. 
Especially noteworthy is the use of wax for fabricating jammable skin chambers, which stiffens by vacuuming \cite{steltz2009soft}. 
The gripper presented in this work also uses a lost-wax molding technique, but with a different type of actuation in mind: the obtained cavity structures are inflated to cause the gripper to bend.

%\rkk{Volder et al. \cite{volder2010pneumatic} provides a detailed review of both pneumatic and hydraulic microactuators.} 
There are several hardware examples for soft grippers described in recent literature, we will mainly focus on fluidic-based systems.
Deimel and Brock \cite{deimel2013compliant} developed a pneumatically actuated three-fingered hand made of reinforced silicone that is mounted to a hard robot and capable of grasping.
More recently, they have developed an anthropomorphic soft pneumatic hand capable of dexterous grasps, which is not mounted to a robot, but instead held by a human \cite{deimel2014novel}. 
Using a soft-lithography fabrication, Ilievski et al. \cite{ilievski2011soft} created a pneumatic starfish-like gripper composed of an array of silicone chambers and a PDMS membrane. 
The gripper hangs on a string and grasps objects like an egg or a mouse in an open-loop controlled manner. 
Stokes et al. \cite{Stokes2014hybrid} use a soft elastomer quadrupedal robot attached to a wheeled robot to grip and retrieve objects. 
A puncture resistant soft pneumatic gripper is developed by Shepherd et al. in \cite{shepherd2013soft}. 
An alternative to positive pressure actuated soft grippers is a robotic gripper that makes use of granular material jamming developed by Brown et al. and detailed in \cite{brown2010universal}.
\rkk{Ikuta and Suzuki \cite{ikuta2002safety} demonstrated a multiple-segment micro-hydraulic actuator for entering blood vessels.}
The soft octopus-inspired arms developed in \cite{calisti2010study} and \cite{calisti2011octopus} are not fluidic powered, but instead use cables to pull rigid fixtures embedded within an elastomer body. 
The arms were capable of grasping objects like pens or screws.
A soft robotic tentacle developed in \cite{martinez2013robotic} was able to hold a flower and a horseshoe-shaped object.
The closest related soft pneumatic actuator design to our current work is the fast Pneu-net designs by Mosadegh et al. \cite{mosadegh2014pneumatic} and by Polygerinos et al. \cite{polygerinos2013towards}.
These finger-like actuators deform with minimal volume change and can bend to high curvatures.
None of the above described grippers were controlled autonomously to perform their tasks and accordingly no statement on repeatability of the autonomous execution was given.

%talk about control
\rkk{Simulation results using an online motion planner for a planar continuum manipulators were presented by Xiao and Vatcha\cite{xiao2010realtime}.}
\rkk{This work was extended by Li and Xiao \cite{li2015general} to present a more general formulation to constrained, continuum manipulation.}
%More generally, the literature has several examples of soft pneumatic elastomer robots. 
%Many of these robots use open-loop controllers. 
%There are soft fluid-powered rolling robots \cite{correll2010soft, onal2011soft, marchese2011soft}. 
%Also, there are soft-bodied robotic fish powered by pneumatics \cite{marchese2014autonomous} and hydraulics \cite{katzschmann2014hydraulic}. 
%There are soft legged pneumatic walkers \cite{shepherd2011multigait, tolley2014resilient}, explosive jumpers \cite{shepherd2013using}, and 
Marchese et al. \cite{marchese2014design} demonstrated closed-loop position control of a multi-segment soft planar fluidic elastomer manipulator in free space. 
%Positioning is achieved using forward and inverse kinematic models and real-time camera measurements. 
%The manipulator curvatures are controlled in real-time using a curvature-volume cascaded control law.
\rkk{The soft manipulator presented in \cite{marchese2014whole} is only suitable for inspection tasks by moving through a constrained environment without object interaction or manipulation.}
%Previously, there were only single markers along the arm, which worked for a limited motion range, but was prone to fail if the markers go out of sight or another object is introduced to the environment. 
%Also, previously many trials were needed to successfully arrive at a final position.

\rkk{The work presented here combines two soft actuator types to leverage their strengths in a planar manipulation task. A new control method described here demonstrates the capabilities of this new soft manipulator to perform autonomous manipulations with uncertainties.}
%The control of the arm is improved from the previous version by using a more robust vision tracking approach.

### 2015-10-16 katzschmann2015autonomous.tex
remove old /rkk comments
all changes to the intro based on daniela's input.
changes to fab section
review comments by David implemented

\newcommand{\unit}[1]{\ensuremath{\, \mathrm{#1}}} %Use this macro to tell \LaTeX that you don’t want your units slanted in math mode.
%\newcommand{\rkk}[1]{{\color{blue}#1}} 
\newcommand{\rkk}[1]{#1}

@@

\title{\LARGE \bf
Autonomous Object Manipulation using a Soft Planar Grasping Manipulator}
\author{Robert K. Katzschmann, Andrew D. Marchese, and Daniela Rus% <-this % stops a space
\thanks{Robert K. Katzschmann, Andrew D. Marchese, and Daniela Rus are with the Computer Science and Artificial Intelligence Laboratory, Massachusetts Institute of Technology, 32 Vassar St. Cambridge, MA 02139, USA, {\tt\small \{rkk, andy, rus\}@csail.mit.edu}}%
}

@@

\maketitle
\thanks
%\twocolumn[
  %\begin{@twocolumnfalse}
    \begin{abstract}
	\textbf{
      	\input{abstract}
	}
    \end{abstract}
    %\bigskip
  %\end{@twocolumnfalse}
%]
%\thispagestyle{empty}
%\pagestyle{plain}

@@

\section*{ACKNOWLEDGMENTS}
This research was conducted in the Distributed Robotics Laboratory at MIT with support from the National Science Foundation, grant numbers NSF 1117178, NSF IIS1226883 and NSF CCF1138967, and National Science Foundation Graduate Research Fellowship Program, primary award number 1122374. We are grateful for this support.

>>>>

\newcommand{\unit}[1]{\ensuremath{\, \mathrm{#1}}} %Use this macro to tell \LaTeX that you don’t want your units slanted in math mode.
\newcommand{\rkk}[1]{{\color{blue}#1}} 
%\newcommand{\rkk}[1]{#1} 
\newcommand{\dr}[1]{{\color{blue}#1}} 
%\newcommand{\dr}[1]{#1}

@@

\title{\LARGE \bf
Autonomous Object Manipulation using a Soft Planar Grasping Manipulator}
\author{Robert K. Katzschmann, Andrew D. Marchese, and Daniela Rus% <-this % stops a space
\thanks{Robert K. Katzschmann, Andrew D. Marchese, and Daniela Rus are with the Computer Science and Artificial Intelligence Laboratory, Massachusetts Institute of Technology, 32 Vassar St. Cambridge, MA 02139, USA, {\tt\small rkk@csail.mit.edu}}%
}

@@

\maketitle
\thanks
%\twocolumn[
  %\begin{@twocolumnfalse}
    \begin{abstract}
\textbf{\input{abstract}}
    \end{abstract}
    %\bigskip
  %\end{@twocolumnfalse}
%]
%\thispagestyle{empty}
%\pagestyle{plain}

@@

\section*{ACKNOWLEDGMENTS}
This research was conducted in the Distributed Robotics Laboratory at MIT with support from the National Science Foundation, grant numbers NSF 1117178, NSF IIS1226883, NSF CCF1138967, and National Science Foundation Graduate Research Fellowship Program, primary award number 1122374. We are grateful for this support.

### 2015-10-16 processing_control.tex
review comments by David implemented

\subsection{Kinematic Control}
The forward kinematics algorithm \texttt{forwKin()} assumes piece-wise constant curvature according to \cite{webster2010design} and \cite{marchese2014design}. 
%This algorithm was experimentally validated for the soft planar arm in \cite{marchese2014design}.
In order to uniquely fit a configuration representation to measured endpoint data in real-time, we use a previously developed single segment inverse kinematics algorithm \texttt{singSegInvKin()} \cite{marchese2014design}.
The inputs to this block are the start and endpoint measurements in $\mathbb{R}^2$: $\mathbf{E}_n \; \forall n  = 1..N$, where $N$ is the number of segments composing the arm.
The outputs from this block are the representations of the measured manipulator configuration: \rkk{measured curvature} $\boldsymbol{\kappa}_{\textrm{meas}}$ and \rkk{segment length} $\mathbf{L}_{\textrm{meas}}$.
%\rkk{Each segment of the manipulator is described by its measured curvature $\kappa_{\textrm{meas}}$ and its segment length $L_{\textrm{meas}}$}.
A cascaded closed-loop curvature controller \texttt{curvatureController()} takes the target curvatures $\boldsymbol{\kappa}_{\textrm{target}}$ as its input and resolves the error between $\boldsymbol{\kappa}_{\textrm{target}}$ and $\boldsymbol{\kappa}_{\textrm{meas}}$ by continuously adjusting the fluidic drive cylinder array.

\rkk{We build on the path planning algorithm presented in \cite{marchese2014whole}. This prior work plans the motion of a soft arm without a gripper through a maze at a centerline while taking the arms bulging shape into account. The approach does not work for approaching and grasping objects, since a tip trajectory for successfully moving towards the object is not known, but needs to be generated by posing and solving a new optimization problem. This tip trajectory has to avoid that the manipulator trunk pushes the object prematurely away when approaching it. Therefore, a new planner had to be developed and is presented in the following section.}

@@

\begin{figure}[!htb]
\centering
   \includegraphics[width=0.95\columnwidth]{Figures/processing_control/grasp_place_planner.eps}
   \caption{State flow diagram of the grasp-and-place planner developed for the autonomous grasp-and-place operation of the manipulator. This diagram describes essentially the flow of information from the motion tracking system to the discrete hardware.}
   \label{fig:grasp-and-place-planner}
\end{figure}

The Grasp Object Planner passes the approach configurations $\boldsymbol{\kappa}_i^*$ of the arm to the \texttt{curvatureController()} for execution in real-time.
The controller receives measured curvatures $\boldsymbol{\kappa}_{meas}$ and lengths $\boldsymbol{L}_{meas}$ at an update rate of $100\unit{Hz}$ from the recursively called \texttt{singSegInvKin()} and uses them to successively control the arm to every intermediate configuration $\boldsymbol{\kappa}_i^*$.
During the arm initialization, the new curvature controller performs a pre-pressurization of both lateral channels. 
This is only done for the two segments closest to the root of the arm in order to stiffen them and shorten their response time constant.  
To allow for smoother transitions between each configuration $\boldsymbol{\kappa}_i^*$, we also added a trajectory generation procedure \texttt{trajGen()} to the new curvature controller. It generates in real-time velocity profiles with acceleration and velocity constraints for each individual degree-of-freedom.
These profiles allow real-time interpolation between the approach configurations of the arm while avoiding overshooting at the next target configuration.
When the arm has arrived at the desired pose next to the object, the curvature controller initiates \texttt{graspObject()}.
After encapsulating the object, \texttt{moveToBin()} requests \texttt{trajGen()} for another trajectory from the current pose to a pre-defined bin location.
\rkk{When the manipulator gets close to the bin location}, the procedure \texttt{releaseObject()} causes the gripper to open and release the object.

@@

\begin{algorithm}[!htbp]
\begin{small}
  \SetAlgoLined\DontPrintSemicolon
  \KwIn{
  $\boldsymbol{\kappa}_{\textrm{meas}}$, $\mathbf{L}_{\textrm{meas}}$ $\gets$ measured arm configuration \;
  $\boldsymbol{\kappa}_{\textrm{off}}$ $\gets$ \rkk{measured manipulator} configuration at start \;
  $g_{\textrm{off}}$ $\gets$ gripper offset normal to end-effector\;
  $x_o$, $y_o$, $r_o$ $\gets$ object center coordinates and radius \;
  $N$ $\gets$ number of manipulator segments}
  \SetKwFunction{planGrasp}{planGrasp}
  \SetKwFunction{findOptimalConfig}{findOptimalConfig}
  \SetKwFunction{forwKin}{forwKin}
  \SetKwProg{myproc}{Procedure}{}{}
  \myproc{\planGrasp{}}{
  $\mathbf{w}_t \gets \forwKin \left( \boldsymbol{\kappa}_{\textrm{meas}}, \, \mathbf{L}_{\textrm{meas}}, \, N, \, L_N \right)$. \;
  $d_1 \gets \| \left[ x_o, \, y_o \right]^{\textrm{T}} \, - \, \mathbf{w}_t \|$. \;
  $d_2 \gets d_1 - r_o - g_{\textrm{off}}$. \;
  $numMoves \gets \lfloor \frac{d_2}{\Delta d} \rfloor$. \;
  $i$ = 0. \;
  \Repeat{$i$ = $numMoves$}{
    $i$ = $i \, + \, 1$ \;
    $r_{a_i} \gets d_1 \, - \, i\,\frac{d_2}{numMoves}$. \;
    $\boldsymbol{\kappa}_i^* \gets$ \findOptimalConfig{$r_{a_i}$}\;
  }
  \KwRet{$\boldsymbol{\kappa}_i^* \quad \forall i \, = 1.. numMoves$} \;}{}
  \setcounter{AlgoLine}{0}
  \SetKwProg{myproc}{Procedure}{}{}
  \myproc{\findOptimalConfig{$r_{a_i}$}}{
  $\begin{aligned}
    & \boldsymbol{\kappa}^* \, \gets \, \underset{\phi, \, \boldsymbol{\kappa}}{\text{min}}
    & & \mathbf{R} \, \left( \boldsymbol{\kappa} \, - \, \boldsymbol{\kappa}_{\textrm{off}} \right)^2. \\
    & \text{subject to}
    & &  \mathbf{w}_t \gets \begin{bmatrix} x_o \, + \, r_{a_i} \, \cos{\phi} \\ y_o \, + \, r_{a_i} \, \sin{\phi} \\ \phi \, + \, \frac{\pi}{2} \end{bmatrix}. \\
    & & & \mathbf{f} \gets \forwKin\left( \boldsymbol{\kappa}, \, \mathbf{L}, \, N, \, L_N \right). \\
    & & & \mathbf{w}_t - \mathbf{w}_{\textrm{off}}\left(r_o, \, \phi \right) - \mathbf{f} \, = \, \mathbf{0}. \\
    & & & \kappa_{n}^{\text{min}} \leq \kappa_n \leq \kappa_{n}^{\text{max}} \quad \forall \, n = 1 \, .. \,N. \\
    \end{aligned}$ \;
  \KwRet{$\boldsymbol{\kappa}^*$} \; }
  \setcounter{AlgoLine}{0}
  \SetKwProg{myproc}{Procedure}{}{}
  \myproc{\forwKin{$\boldsymbol{\kappa}$, $\mathbf{L}$, $i$, $s$ }}{
  \KwIn{$\boldsymbol{\kappa}$, $\mathbf{L}$, $i$ the segment of interest index, $s$ the arc length along the indexed segment}
  \eIf{$i = 0$}{
        $\theta_{i}(0) \gets \theta_0(0)$. \;
        $x_{i}(0) \gets 0$.\;
        $y_{i}(0) \gets 0$.\; }
    {
        $\left[ x_{i}(0), \, y_{i}(0), \, \theta_{i}(0) \right] \gets \forwKin \left(\boldsymbol{\kappa}, \, \mathbf{L}, \, i-1, \, L_{i-1} \right)$.\;
    }
    $\theta \gets \theta_{i}(0) + k_i s$. \;
    $x \gets x_{i}(0) + \frac{\sin{\theta}}{k_i} - \frac{\sin{\theta_{i}(0)}}{k_i}$. \;
    $y \gets y_{i}(0) - \frac{\cos{\theta}}{k_i} + \frac{\cos{\theta_{i}(0)}}{k_i}$. \;
  \KwRet{$\left[ x, \, y, \, \theta \right]^{\mathrm{T}}$ or $\left[ x, \, y \right]^{\mathrm{T}}$}}
  \caption{Grasp Object Planner}
  \label{alg:planGrasp}
\end{small}

@@

\rkk{The nonlinear optimization problem was implemented using Matlab’s Optimization Toolbox with the function calls \emph{fmincon}, which finds the minimum of a constrained nonlinear multivariable function. Sequential Quadratic Programming was used as the solver with a relative upper bound of $2 \times 10^{-3}$ on the magnitude of the constraint functions. The lower bound on the size of a step was given by $1 \times 10^{-6}$. The solver takes about 1\unit{s} on a regular PC to solve for all waypoints from start to finish.}

>>>>

\subsection{Kinematic Control}
The forward kinematics algorithm \texttt{forwKin()} assumes piece-wise constant curvature \cite{webster2010design}. 
%This algorithm was experimentally validated for the soft planar arm in \cite{marchese2014design}.
In order to uniquely fit a configuration representation to measured endpoint data in real-time, we use a previously developed single segment inverse kinematics algorithm \texttt{singSegInvKin()} \cite{marchese2014design}.
The inputs to this block are the start and endpoint measurements in $\mathbb{R}^2$: $\mathbf{E}_n \; \forall n  = 1..N$, where $N$ is the number of segments composing the arm.
The outputs from this block are the representations of the measured manipulator configuration: \rkk{measured curvature} $\boldsymbol{\kappa}_{\textrm{meas}}$ and \rkk{segment length} $\mathbf{L}_{\textrm{meas}}$.
%\rkk{Each segment of the manipulator is described by its measured curvature $\kappa_{\textrm{meas}}$ and its segment length $L_{\textrm{meas}}$}.
A cascaded closed-loop curvature controller \texttt{curvatureController()} takes the target curvatures $\boldsymbol{\kappa}_{\textrm{target}}$ as its input and resolves the error between $\boldsymbol{\kappa}_{\textrm{target}}$ and $\boldsymbol{\kappa}_{\textrm{meas}}$ by continuously adjusting the fluidic drive cylinder array.

\rkk{We build on the path planning algorithm presented in \cite{marchese2014whole}. This prior work plans the motion of a soft arm without a gripper through a maze at a centerline while taking the arms bulging shape into account. The approach does not work for approaching and grasping objects, since a tip trajectory for successfully moving towards the object is not known, but needs to be generated by posing and solving a new optimization problem. The manipulator trunk should not push the object away when approaching it. Therefore, a new planner had to be developed and is presented in the following section.}

@@

\begin{figure}[!htb]
\centering
   \includegraphics[width=0.95\columnwidth]{Figures/processing_control/grasp_place_planner.eps}
   \caption{State flow diagram of the planner developed for the autonomous grasp-and-place operation of the manipulator. This diagram describes essentially the flow of information from the motion tracking system to the discrete hardware.}
   \label{fig:grasp-and-place-planner}
\end{figure}

The Grasp Object Planner passes the approach configurations $\boldsymbol{\kappa}_i^*$ of the arm to the \texttt{curvatureController()} for execution in real-time.
The controller receives measured curvatures $\boldsymbol{\kappa}_{\textrm{meas}}$ and lengths $\mathbf{L}_{meas}$ at an update rate of $100\unit{Hz}$ from the recursively called \texttt{singSegInvKin()} and uses them to successively control the arm to every intermediate configuration $\boldsymbol{\kappa}_i^*$.
During the arm initialization, the new curvature controller performs a pre-pressurization of both lateral channels. 
This is only done for the two segments closest to the root of the arm in order to stiffen them and shorten their response time constant.  
To allow for smoother transitions between each configuration $\boldsymbol{\kappa}_i^*$, we also added a trajectory generation procedure \texttt{trajGen()} to the new curvature controller. It generates in real-time velocity profiles with acceleration and velocity constraints for each individual degree-of-freedom.
These profiles allow real-time interpolation between the approach configurations of the arm while avoiding overshooting at the next target configuration.
When the arm has arrived at the desired pose next to the object, the curvature controller initiates \texttt{graspObject()}.
After encapsulating the object, \texttt{moveToBin()} requests \texttt{trajGen()} for another trajectory from the current pose to a pre-defined bin location.
\rkk{When the manipulator gets close to the bin location}, the procedure \texttt{releaseObject()} causes the gripper to open and release the object.

@@

\begin{algorithm}[!htbp]
\begin{small}
  \SetAlgoLined\DontPrintSemicolon
  \KwIn{
  $\boldsymbol{\kappa}_{\textrm{meas}}$, $\mathbf{L}_{\textrm{meas}}$ $\gets$ measured arm configuration \;
  $\boldsymbol{\kappa}_{\textrm{off}}$ $\gets$ \rkk{measured manipulator} configuration at start \;
  $g_{\textrm{off}}$ $\gets$ gripper offset normal to end-effector\;
  $x_o$, $y_o$, $r_o$ $\gets$ object center coordinates and radius \;
  $N$ $\gets$ number of manipulator segments}
  \SetKwFunction{planGrasp}{planGrasp}
  \SetKwFunction{findOptimalConfig}{findOptimalConfig}
  \SetKwFunction{forwKin}{forwKin}
  \SetKwProg{myproc}{Procedure}{}{}
  \myproc{\planGrasp{}}{
  $\mathbf{w}_t \gets \forwKin \left( \boldsymbol{\kappa}_{\textrm{meas}}, \, \mathbf{L}_{\textrm{meas}}, \, N, \, L_N \right)$. \;
  $d_1 \gets \| \left[ x_o, \, y_o \right]^{\textrm{T}} \, - \, \mathbf{w}_t \|$. \;
  $d_2 \gets d_1 - r_o - g_{\textrm{off}}$. \;
  $numMoves \gets \lfloor \frac{d_2}{\Delta d} \rfloor$. \;
  $i$ = 0. \;
  \Repeat{$i$ = $numMoves$}{
    $i$ = $i \, + \, 1$. \;
    $r_{a_i} \gets d_1 \, - \, i\,\frac{d_2}{numMoves}$. \;
    $\boldsymbol{\kappa}_i^* \gets$ \findOptimalConfig{$r_{a_i}$}.\;
  }
  \KwRet{$\boldsymbol{\kappa}_i^* \quad \forall i \, = 1.. numMoves$} \;}{}
  \setcounter{AlgoLine}{0}
  \SetKwProg{myproc}{Procedure}{}{}
  \myproc{\findOptimalConfig{$r_{a_i}$}}{
  $\begin{aligned}
    & \boldsymbol{\kappa}^* \, \gets \, \underset{\phi, \, \boldsymbol{\kappa}}{\text{min}}
    & & \mathbf{R} \, \left( \boldsymbol{\kappa} \, - \, \boldsymbol{\kappa}_{\textrm{off}} \right)^2. \\
    & \text{subject to}
    & &  \mathbf{w}_t \gets \begin{bmatrix} x_o \, + \, r_{a_i} \, \cos{\phi} \\ y_o \, + \, r_{a_i} \, \sin{\phi} \\ \phi \, + \, \frac{\pi}{2} \end{bmatrix}. \\
    & & & \mathbf{f} \gets \forwKin\left( \boldsymbol{\kappa}, \, \mathbf{L}, \, N, \, L_N \right). \\
    & & & \mathbf{w}_t - \mathbf{w}_{\textrm{off}}\left(r_o, \, \phi \right) - \mathbf{f} \, = \, \mathbf{0}. \\
    & & & \kappa_{n}^{\text{min}} \leq \kappa_n \leq \kappa_{n}^{\text{max}} \quad \forall \, n = 1 \, .. \,N. \\
    \end{aligned}$ \;
  \KwRet{$\boldsymbol{\kappa}^*$} \; }
  \setcounter{AlgoLine}{0}
  \SetKwProg{myproc}{Procedure}{}{}
  \myproc{\forwKin{$\boldsymbol{\kappa}$, $\mathbf{L}$, $i$, $s$ }}{
  \KwIn{$\boldsymbol{\kappa}$, $\mathbf{L}$, $i$ the segment of interest index, $s$ the arc length along the indexed segment}
  \eIf{$i = 0$}{
        $\theta_{i}(0) \gets \theta_0(0)$. \;
        $x_{i}(0) \gets 0$.\;
        $y_{i}(0) \gets 0$.\; }
    {
        $\left[ x_{i}(0), \, y_{i}(0), \, \theta_{i}(0) \right] \gets \forwKin \left(\boldsymbol{\kappa}, \, \mathbf{L}, \, i-1, \, L_{i-1} \right)$.\;
    }
    $\theta \gets \theta_{i}(0) + k_i s$. \;
    $x \gets x_{i}(0) + \frac{\sin{\theta}}{k_i} - \frac{\sin{\theta_{i}(0)}}{k_i}$. \;
    $y \gets y_{i}(0) - \frac{\cos{\theta}}{k_i} + \frac{\cos{\theta_{i}(0)}}{k_i}$. \;
  \KwRet{$\left[ x, \, y, \, \theta \right]^{\mathrm{T}}$ or $\left[ x, \, y \right]^{\mathrm{T}}$}}
  \caption{Grasp Object Planner}
  \label{alg:planGrasp}
\end{small}

@@

\rkk{The nonlinear optimization problem is implemented on a PC using Sequential Quadratic Programming, which finds iteratively the minimum of a constrained nonlinear multivariable function. The solver is run with a relative upper bound of $2 \times 10^{-3}$ on the magnitude of the constraint functions. The lower bound on the size of a step was given by $1 \times 10^{-6}$. The solver takes about 1\unit{s} to solve for all waypoints from start to finish.}
